---
title: "Lung Cancer"
output: 
  flexdashboard::flex_dashboard:
   vertical_layout: fill
   orientation: columns
runtime: shiny
---
    
```{r global, include = FALSE}
#refer: install.packages("/Users/helenhoang/Downloads/terra_1.7-55.tgz", repos = NULL, type = "source")

library(flexdashboard)
library(ggplot2)
library(gplots)
library(dplyr)
library(usmap)
library(collapsibleTree)
library(DT)
library(terra)
library(raster)
library(leaflet)

#General Page
file_path1 = "/Users/helenhoang/Downloads/cancer_incidence/lung_cancer/lungcancer_inc_per100k_pop_2015_2019.csv"
lung = read.csv(file_path1)
lungcp <- lung
lungcp$Value <- as.numeric(ifelse(lungcp$Value == "Suppressed", NA, lungcp$Value))
lungcp <- dplyr::select(lungcp, -Start.Year)
lungcp <- dplyr::select(lungcp, -End.Year)
lungcp <- dplyr::select(lungcp, -Data.Comment)
lungcp <- dplyr::select(lungcp, -X)
lungcp <- lungcp %>% rename("fips" = "CountyFIPS")

#Gender Page
file_path2 = "/Users/helenhoang/Downloads/cancer_incidence/lung_cancer/lungcancer_inc_per100k_pop_2015_2019_gender.csv"
lung_gender = read.csv(file_path2)
lung_gendercp <- lung_gender
lung_gendercp$Value <- as.numeric(ifelse(lung_gendercp$Value == "Suppressed", NA, lung_gendercp$Value))
lung_gendercp <- lung_gendercp %>%
  group_by(State, Gender) %>%
  mutate(Value = if_else(is.na(Value), round(mean(Value, na.rm = TRUE), 1), Value))
lung_gendercp <- dplyr::select(lung_gendercp, -Start.Year)
lung_gendercp <- dplyr::select(lung_gendercp, -End.Year)
lung_gendercp <- dplyr::select(lung_gendercp, -Data.Comment)
lung_gendercp <- lung_gendercp %>% rename("fips" = "CountyFIPS")

#Race Page
file_path3 = "/Users/helenhoang/Downloads/cancer_incidence/lung_cancer/lungcancer_inc_per100k_pop_2015_2019_race.csv"
lung_race = read.csv(file_path3)
lung_racecp <- lung_race
lung_racecp$Value <- as.numeric(ifelse(lung_racecp$Value == "Suppressed", NA, lung_racecp$Value))
lung_racecp <- lung_racecp %>%
  group_by(State, Race.Ethnicity) %>%
  mutate(Value = if_else(is.na(Value), round(mean(Value, na.rm = TRUE), 1), Value))
lung_racecp <- dplyr::select(lung_racecp, -Start.Year)
lung_racecp <- dplyr::select(lung_racecp, -End.Year)
lung_racecp <- dplyr::select(lung_racecp, -Data.Comment)
lung_racecp <- lung_racecp %>% rename("fips" = "CountyFIPS")

#Factors

#1.air quality
airquality_path = "/Users/helenhoang/Downloads/OneDrive_1_11-21-2023/air_quality_pm2.5annualavg_bycounty_2018_2019.csv"
airquality = read.csv(airquality_path)

#2.arsenic
arsenic_path = "/Users/helenhoang/Downloads/OneDrive_1_11-21-2023/arsenic_annual_mean_conc_2018_2019.csv"
arsenic = read.csv(arsenic_path)

#3.binge
binge_path = "/Users/helenhoang/Downloads/OneDrive_1_11-21-2023/binge_drinking_alcohol_adults_per100k_pop_2018_2019.csv"
binge = read.csv(binge_path)

#4.coronary
coronary_path = "/Users/helenhoang/Downloads/OneDrive_1_11-21-2023/coronary_heart_disease_per100k_pop_2018_2019.csv"
coronary = read.csv(coronary_path)

#5.asthma
asthma_path = "/Users/helenhoang/Downloads/OneDrive_1_11-21-2023/county_asthma_per100k_pop_2018_2019.csv"
asthma = read.csv(asthma_path)

#6.noHealthIns
noHealthIns_path = "/Users/helenhoang/Downloads/OneDrive_1_11-21-2023/county_noHealthIns_perc_pop_2018_2019.csv"
noHealthIns = read.csv(noHealthIns_path)

#7.diabetes
diabetes_path = "/Users/helenhoang/Downloads/OneDrive_1_11-21-2023/diabetes_adults_per100k_pop_2018_2019.csv"
diabetes = read.csv(diabetes_path)

#8.obesity
obesity_path = "/Users/helenhoang/Downloads/OneDrive_1_11-21-2023/obesity_adults_per100k_pop_2018_2019.csv"
obesity = read.csv(obesity_path)

#9.below_poverty
below_poverty_path = "/Users/helenhoang/Downloads/OneDrive_1_11-21-2023/pop_perc_below_poverty_2018_2019.csv"
below_poverty = read.csv(below_poverty_path)

#10.smoking
smoking_path = "/Users/helenhoang/Downloads/OneDrive_1_11-21-2023/smoking_adults_per100k_pop_2018_2019.csv"
smoking = read.csv(smoking_path)

airquality_avg_value <- airquality %>%
  group_by(State, CountyFIPS, County) %>%
  summarize(Avg_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  rename("fips" = "CountyFIPS") %>%
  rename("Airquality" = "Avg_Value")

arsenic_avg_value <- arsenic %>%
  group_by(State, CountyFIPS, County) %>%
  summarize(Avg_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  rename("fips" = "CountyFIPS") %>%
  rename("Arsenic" = "Avg_Value")

binge_avg_value <- binge %>%
  group_by(State, CountyFIPS, County) %>%
  summarize(Avg_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  rename("fips" = "CountyFIPS") %>%
  rename("Binge" = "Avg_Value")

coronary_avg_value <- coronary %>%
  group_by(State, CountyFIPS, County) %>%
  summarize(Avg_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  rename("fips" = "CountyFIPS") %>%
  rename("Coronary" = "Avg_Value")

asthma_avg_value <- asthma %>%
  group_by(State, CountyFIPS, County) %>%
  summarize(Avg_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  rename("fips" = "CountyFIPS") %>%
  rename("Asthma" = "Avg_Value")

noHealthIns_avg_value <- noHealthIns %>%
  group_by(State, CountyFIPS, County) %>%
  summarize(Avg_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  rename("fips" = "CountyFIPS") %>%
  rename("NoHealthIns" = "Avg_Value")

diabetes_avg_value <- diabetes %>%
  group_by(State, CountyFIPS, County) %>%
  summarize(Avg_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  rename("fips" = "CountyFIPS") %>%
  rename("Diabetes" = "Avg_Value")

obesity_avg_value <- obesity %>%
  group_by(State, CountyFIPS, County) %>%
  summarize(Avg_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  rename("fips" = "CountyFIPS") %>%
  rename("Obesity" = "Avg_Value")

below_poverty_avg_value <- below_poverty %>%
  group_by(State, CountyFIPS, County) %>%
  summarize(Avg_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  rename("fips" = "CountyFIPS") %>%
  rename("Below_poverty" = "Avg_Value")

smoking_avg_value <- smoking %>%
  group_by(State, CountyFIPS, County) %>%
  summarize(Avg_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  rename("fips" = "CountyFIPS") %>%
  rename("Smoking" = "Avg_Value")

all_factors_lung <- lungcp %>%
  left_join(airquality_avg_value, by = c("State", "fips", "County")) %>%
  left_join(arsenic_avg_value, by = c("State", "fips", "County")) %>%
  left_join(binge_avg_value, by = c("State", "fips", "County")) %>%
  left_join(coronary_avg_value, by = c("State", "fips", "County")) %>%
  left_join(asthma_avg_value, by = c("State", "fips", "County")) %>% 
  left_join(noHealthIns_avg_value, by = c("State", "fips", "County")) %>% 
  left_join(diabetes_avg_value, by = c("State", "fips", "County")) %>% 
  left_join(obesity_avg_value, by = c("State", "fips", "County")) %>% 
  left_join(below_poverty_avg_value, by = c("State", "fips", "County")) %>% 
  left_join(smoking_avg_value, by = c("State", "fips", "County")) %>%
  rename("Lung_Cancer" = "Value") 

all_factors_lung
```

General {data-icon=""}
===================================== 

Column {data-width=600 .tabset .tabset-fade}
-------------------------------------

### Filter by State

```{r}
names(obesity)
```


#### Filter 
```{r}
options <- unique(lungcp$State)
selectInput("state", label = "Choose custom state:",
            choices = options, selected = "Alabama")
```

#### Map 
```{r}
#plot_usmap(regions = "county", data=lungcp, values = "Value")
renderPlot({
  plot_usmap(regions = "county", 
             data=lungcp, values = "Value", 
             include = input$state, 
             labels = TRUE) + scale_fill_continuous(low = "white", high = "red", name = "Rate of Cancer", label = scales::comma) + labs(title = input$state, subtitle = "This is a rate of lung cancer of the counties") + theme(panel.background=element_blank())
},
height = 500, 
width = 600,
)
```


### Filter by Connty
```{r}
#Source:
# Get USA polygon data
USA <- getData("GADM", country = "usa", level = 2)
# Prepare data
mydata <- all_factors_lung %>% dplyr::select(State, County, Lung_Cancer, 
                                      Airquality, Arsenic, Binge, Coronary, Asthma, 
                                      NoHealthIns, Diabetes, Obesity, Below_poverty, Smoking)
#head(mydata)

temp <- merge(USA, mydata,
              by.x = c("NAME_1", "NAME_2"), by.y = c("State", "County"),
              all.x = TRUE)
# Create a color palette
mypal <- colorNumeric(palette = "viridis", domain = temp$Lung_Cancer, na.color = "grey")

leaflet() %>% 
addProviderTiles("OpenStreetMap.Mapnik") %>%
setView(lat = 39.8283, lng = -98.5795, zoom = 4) %>%
  addPolygons(data = temp, stroke = FALSE, smoothFactor = 0.2, 
              fillOpacity = 0.7,
              fillColor = ~mypal(Lung_Cancer),
              color = "white",
              popup = paste("State: ", temp$NAME_1, "<br>",
                            "County: ", temp$NAME_2, "<br>",
                            "Rate of Lung Cancer: ", temp$Lung_Cancer, "<br>",
                            "Airquality: ", temp$Airquality, "<br>",
                            "Arsenic: ", temp$Arsenic, "<br>",
                            "Binge: ", temp$Binge, "<br>",
                            "Coronary: ", temp$Coronary, "<br>",
                            "Asthma: ", temp$Asthma, "<br>",
                            "NoHealthIns: ", temp$NoHealthIns, "<br>",
                            "Diabetes: ", temp$Diabetes, "<br>",
                            "Obesity: ", temp$Obesity, "<br>",
                            "Below_poverty: ", temp$Below_poverty, "<br>",
                            "Smoking: ", temp$Smoking, "<br>"),
              highlight = highlightOptions(color = "black", bringToFront = TRUE)) %>%
  addLegend(position = "bottomleft", pal = mypal, values = temp$Lung_Cancer,
            title = "Rate of Cancer",
            opacity = 0.7)

```

Column {data-width=400 .tabset .tabset-fade}
-------------------------------------

### Table 

```{r}
DT::datatable(lungcp[, c("State", "County", "Value")],
              rownames = FALSE, options = list(pageLength = 10))
```   

### Analysis
   
#### Top 5th States which have the highest rate of lung cancer
```{r}
top_states <- lungcp %>%
  group_by(State) %>%
  summarise(Avg_Value = mean(Value, na.rm = TRUE)) %>%
  arrange(desc(Avg_Value)) %>%
  head(5)

# Print the top states
print(top_states)
```

#### Top 5th States which have the highest rate of lung cancer
```{r}
top_counties <- lungcp %>%
  group_by(State, County) %>%
  summarise(Avg_Value = mean(Value, na.rm = TRUE)) %>%
  arrange(desc(Avg_Value)) %>%
  head(5)

# Print the top counties
print(top_counties)
```

Gender {data-icon="transgender-outline"}
===================================== 

Column {data-width=500}
-------------------------------------

### Filter by Gender {data-height=250}
   
```{r}
options <- unique(lung_gendercp$Gender)
selectInput("gender", label = "Choose custom gender:",
            choices = c("All", "Female", "Male"), selected = "All")
```


### US Map {data-height=500}
   
```{r}
renderPlot({
  if (input$gender == "All") {
    subset_map_gender <- lung_gendercp
  } else {
    # Subset data for the specific race/ethnicity
    subset_map_gender <- lung_gendercp[lung_gendercp$Gender == input$gender, ]
  }
  plot_usmap(regions = "county", data=subset_map_gender, values = "Value")
})
```


Column {data-width=500 .tabset .tabset-fade}
-------------------------------------
   
### Table

```{r}
renderDT({
  if (input$gender == "All") {
    subset_table_gender <- lung_gendercp
  } else {
    # Subset data for the specific race/ethnicity
    subset_table_gender <- lung_gendercp[lung_gendercp$Gender == input$gender, ]
  }
  DT::datatable(subset_table_gender[, c("State", "County", "Value", "Gender")], 
                  rownames = FALSE, options = list(pageLength = 10))
})
```

### Avg Rate

```{r}
ggplot(lung_gendercp, aes(x = Gender, y = Value, fill = Gender)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge", width = 0.5, height = 0.5) +
  labs(title = "Average Rate of Lung Cancer by Gender", x = "Gender", y = "Average Value") +
  theme_minimal()
```

### Facts

#### State and county with the maximum rate of lung cancer by gender
```{r}
max_value_rows <- lung_gendercp %>%
  group_by(Gender) %>%
  filter(Value == max(Value))

for (i in seq_len(nrow(max_value_rows))) {
  cat("Gender:", max_value_rows$Gender[i], "\n",
      "State with the maximum value:", max_value_rows$State[i], "\n",
      "County with the maximum value:", max_value_rows$County[i], "\n",
      "Value:", max_value_rows$Value[i], "\n\n"
  )
}
```

#### County where the sum of female values is greater than the sum of male values 
```{r}
county_female_greater_than_male <- lung_gendercp %>%
  group_by(State, County) %>%
  summarize(Female = Value[Gender == "Female"],
            Male = Value[Gender == "Male"],
            .groups = "drop") %>%
  filter(Female > Male) %>%
  dplyr::select(State, County, Female, Male)

print(county_female_greater_than_male)
```

Race {data-icon="people-outline"}
===================================== 

Column {data-width=500}
-------------------------------------

### Filter by Race {data-height=250}
   
```{r}
options <- unique(lung_racecp$Race.Ethnicity)
selectInput("race", label = "Choose custom race:",
            choices = c("All", "Asian/Pacific Islander (includes Hispanic)", "Black (includes Hispanic)", "American Indian/Alaskan Native (includes Hispanic)", "White (includes Hispanic)"), selected = "All")
```

### US Map {data-height=500}
   
```{r}
renderPlot({
  if (input$race == "All") {
    subset_map_race <- lung_racecp
  } else {
    # Subset data for the specific race/ethnicity
    subset_map_race <- lung_racecp[lung_racecp$Race.Ethnicity == input$race, ]
  }
  plot_usmap(regions = "county", data=subset_map_race, values = "Value")
})
```

Column {data-width=500 .tabset .tabset-fade}
-------------------------------------
   
### Table

```{r}
#DT::datatable(lung_racecp[, c("State", "County", "Value", "Race.Ethnicity")],
#              rownames = FALSE, options = list(pageLength = 10))
renderDT({
  if (input$race == "All") {
    subset_table_race <- lung_racecp
  } else {
    # Subset data for the specific race/ethnicity
    subset_table_race <- lung_racecp[lung_racecp$Race.Ethnicity == input$race, ]
  }
  DT::datatable(subset_table_race[, c("State", "County", "Value", "Race.Ethnicity")], 
                  rownames = FALSE, options = list(pageLength = 10))
})
```
### Avg Rate

```{r}
# Round only the non-infinite values in the "Value" column
lung_racecp$Value[is.finite(lung_racecp$Value)] <- round(lung_racecp$Value[is.finite(lung_racecp$Value)], 2)

ggplot(lung_racecp, aes(x = Race.Ethnicity, y = Value, fill = Race.Ethnicity)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  labs(title = "Average Rate of Lung Cancer by Race/Ethnicity", x = "Race", y = "Average Value") +
  theme_minimal() +
  theme(axis.text.x = element_blank())  # Hide x-axis labels
```

### Facts

#### State and county with the maximum rate of lung cancer by race
```{r}
max_value_rows_race <- lung_racecp %>%
  filter(!is.na(Race.Ethnicity)) %>%
  group_by(Race.Ethnicity) %>%
  filter(Value == max(Value, na.rm = TRUE))

for (i in seq_len(nrow(max_value_rows_race))) {
  cat("Race.Ethnicity:", max_value_rows_race$Race.Ethnicity[i], "\n",
      "State with the maximum value:", max_value_rows_race$State[i], "\n",
      "County with the maximum value:", max_value_rows_race$County[i], "\n",
      "Value:", max_value_rows_race$Value[i], "\n\n"
  )
}
```

Factors {data-icon=""}
===================================== 

Column {data-width=500}
-------------------------------------

### Filter {data-height=250}
   
```{r}
selectInput("factor", label = "Choose custom factor:",
            choices = c("Airquality", "Arsenic", "Binge", "Coronary", "Asthma", "NoHealthIns", "Diabetes", "Obesity", "Below_poverty", "Smoking"), selected = "Airquality")
```

### Factor {data-height=500}
   
```{r}
renderPlot({
  plot_usmap(regions = "county", data=all_factors_lung, values = input$factor)
})
```

Column {data-width=500 .tabset .tabset-fade}
-------------------------------------

### Table   

```{r}
renderDT({
  DT::datatable(all_factors_lung[, c("State", "County", input$factor)],
    rownames = FALSE, options = list(pageLength = 10)
  )
})
```

### Figures

##### Mean
```{r}
renderText({
  target_factor <- input$factor
  average_value <- mean(all_factors_lung[[target_factor]], na.rm = TRUE)

  # Print the information
  paste("Average of ", target_factor, ":", average_value)
})
```

##### Max
```{r}
renderText({
  target_factor <- input$factor
  max_value <- max(all_factors_lung[[target_factor]], na.rm = TRUE)
  
  # Print the information
  paste("Max of ", target_factor, ":", max_value)
})
```

##### Missing Values
```{r}
renderText({
  target_factor <- input$factor
  missing_values <- sum(is.na(all_factors_lung[[target_factor]]))
  total_rows <- nrow(all_factors_lung)
  
  # Print the information
  paste("Missing Values (marked as NA) of ", target_factor, ":", missing_values, "/", total_rows, "in total")
})
```

##### Top 5th States & Counties which have the highest influence by:
```{r}
renderDataTable({
  target_factor <- input$factor
  # Find the top 5 rows with the highest values in the target_factor column
  top_5_rows <- head(all_factors_lung[order(-all_factors_lung[[target_factor]]), ], 5)
  datatable(top_5_rows[, c("State", "County", input$factor)], 
            options = list(pageLength = 5))
})
```

### Correlations 
    
#### Heat Map
```{r}
cl <- all_factors_lung[, c(
  "Lung_Cancer",
  "Airquality",
  "Arsenic",
  "Binge",
  "Coronary",
  "Asthma",
  "NoHealthIns",
  "Diabetes",
  "Obesity",
  "Below_poverty",
  "Smoking"
)]

cl <- apply(cl, 2, function(x) ifelse(is.na(x), mean(x, na.rm = TRUE), x))

# Compute the correlation matrix
corr <- cor(cl)

# Create a heatmap of the correlation matrix without hierarchical clustering
heatmap.2(corr, 
          dendrogram = 'none', 
          col = colorRampPalette(c("skyblue", "white", "pink"))(50), 
          scale = 'none',
          keysize = 1.5,
          trace = 'none',
          main = 'Heatmap of Correlation Matrix',
          Rowv = FALSE,
          Colv = FALSE,
          cexRow = 0.5,
          cexCol = 0.5)
```

#### Extract the correlation of each variable with lung cancer rate
```{r}
corr_deposite <- data.frame(variable = rownames(corr), corr = corr[, 'Lung_Cancer']) %>%
  filter(variable != 'Lung_Cancer') %>%
  arrange(desc(corr))
corr_deposite
#DT::datatable(corr_deposite[, c("variable", "corr")],  rownames = FALSE, options = list(pageLength = 10))
```

### Summary

```{r}
summary(all_factors_lung)
```

